The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-11 writing/draft3/draft11.tex
witing results

Besides our main model, we estimated the same specification for two additional subsamples of Kenyan counties. In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate. The estimates of all three models are summarised in table~\ref{MainEst}. The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties. All the predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance. As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance in table~\ref{MainEst}. The F-tests were conducted using the marginal rather than sequential sum of squares. That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable. In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative rather than additive. As in the classical linear regression model, a negative sign of a coefficient implies a negative effect and positive sign implies a positive effect. The estimates of the exponents of the coefficients, which can be interpreted as
additive marginal effects on yield are in table~\ref{KenARe11_exponents} in the appendix.

We will now focus on the estimates for the sample of all counties, that is the first two columns in table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yields while average temperature has negative effects on yields for the case of all counties.  We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,

@@

{
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficients \\ Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11a  KEN11a_ASAL    KEN11a_nonASAL										
\label{KenARe11_exponents} 
\begin{footnotesize}
\lineup
\begin{indented}
\item[]\begin{tabular}{@{}llrlllr} 
\br   &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&F-value\tnote{a}% anova(KEN11a, type='marginal')
    &{$exp(\beta)$}&F-value\tnote{a}&{$exp(\beta)$}&F-value\tnote{a}\\
\mr
\\
\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$\05.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$\04.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$\03.277$&$0.969$ &$0.246$&$0.909^{}$&$\02.231$\\
  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$\04.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$\00.163$\\
  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$\04.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$\00.096$\\
  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$\01.580$\\
  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$\03.125$&$1.039$&$0.558$&$1.059$ ${*}$&$\05.640$\\
  \\
  \hline\\[-1em]
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
  \\[-1em]
\hline
\\[-1em]Intercept\\
 \\[-1em] \hline
\\[-1em]
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}
\\
\br
\end{tabular} 
\end{indented}
\end{footnotesize}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
        \begin{adjustwidth}{1cm}{} 
    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

>>>>

Besides our main model, we estimated the same specification for two additional subsamples of Kenyan counties. In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate. The estimates of all three models are summarised in table~\ref{MainEst}. The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties. All predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance.

As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance in table~\ref{MainEst}. The F-tests were conducted using the marginal rather than sequential sum of squares. That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable. In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the effects on yield are multiplicative. The exponents of the estimates (which can be interpreted as additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.

First, we will focus on the estimates based on the sample of all counties, that is the first two columns in table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yield while average temperature has negative effects on yields.

We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,

@@

{
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficient estimates\\ Log of maize yield and weather, ARMA(1,1) errors}}
%KEN11a  KEN11a_ASAL    KEN11a_nonASAL										
\label{KenARe11_exponents} 
\begin{footnotesize}
\lineup
\begin{indented}
\item[]\begin{tabular}{@{}llrlllr} 
\br   &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&F-value\tnote{a}% anova(KEN11a, type='marginal')
    &{$exp(\beta)$}&F-value\tnote{a}&{$exp(\beta)$}&F-value\tnote{a}\\
\mr
\\
\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$\05.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$\04.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$\03.277$&$0.969$ &$0.246$&$0.909^{}$&$\02.231$\\
  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$\04.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$\00.163$\\
  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$\04.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$\00.096$\\
  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$\01.580$\\
  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$\03.125$&$1.039$&$0.558$&$1.059$ ${*}$&$\05.640$\\
  \\
  \hline\\[-1em]
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
  \\[-1em]
\hline
\\[-1em]Intercept\\
 \\[-1em] \hline
\\[-1em]
\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}
\\
\br
\end{tabular} 
\end{indented}
\end{footnotesize}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
        \begin{adjustwidth}{1cm}{} 
    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
